Open archaeology, open source?

### 2023-09-20 analysis/_01-introduction.qmd
Merge outline into text
Modifications and comments.

@@

However, academic open source has a complicated relationship with open source as practiced by professional software developers, which has its own distinct history and is framed by different objectives, challenges, and value regimes. Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source. For instance, the Open Knowledge Foundation [-@openknowledgefoundation2023] publishes a widely accepted defitnion of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group. The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic). However, academic work, including the development of academic software, differs signififcantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL, and the Firefox web browser. While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests. In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@kelty2008 xx].

>>>>

<!--
* There is open archaeology
* Open archaeology is related to open source
* Is open archaeology open source, and is open source open?
  * [Is 'open archaeology' just uploading CSVs?]
-->

@@

However, academic open source has a complicated relationship with open source as practiced by professional software developers, which has its own distinct history and is framed by different objectives, challenges, and value regimes. Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source. For instance, the Open Knowledge Foundation [-@openknowledgefoundation2023] publishes a widely accepted definition of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group. The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plain text, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic). However, academic work, including the development of academic software, differs significantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL, and the Firefox web browser. While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests. In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@kelty2008 xx].

### 2023-09-20 analysis/_02-background.qmd
Merge outline into text
Modifications and comments.

# Open science and open source <!-- Background -->

@@

Open source has traditionally been referred to as being based on meritocratic principles. A good test of whether a contribution should be included in a published software release is whether it is functional [@kelty2008 220]. Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues. This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion. Personal ego is minimized in favour of co-creating stable and functional outcomes [@coleman2012; @oneil2009 on the debian saga].

@@

In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly underappreciated factors that contribute to an open source project's success.

@@

The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient. Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" [@openknowledgefoundation2023]. This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement. However, most open science advocates (including archaeologists, as elicited by @beck2012 and @marwick2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.

The idea that scientistis should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals [@mirowski2018]. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes. However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain [@tennant2020]. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (inlcuding: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out. Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software.

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards [cf. @kansa2014; @kintigh2015; @roosevelt2015] distract from fundamental tensions and contradictions regarding the actual value of working in the open. For instance, Faniel et al. [-@faniel2013 299-301], Atici et al. [-@atici2013 676-677], @huggett2018, @sobotkova2018, and @opitz2021 demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate. In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation. Here we aim to identify similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists.

>>>>

# Open science and open source

<!--
* What is open source?
* Summary of key publications in open archaeology / DIY archaeology
* But is any of this real? There seems to be a lack of reflection about what open archaeology is.
  -->

@@

Open source has traditionally been referred to as being based on meritocratic principles. A good test of whether a contribution should be included in a published software release is whether it is functional [@kelty2008 220]. Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues. This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion. Personal ego is minimized in favour of co-creating stable and functional outcomes [@coleman2012; @oneil2009 on the Debian saga].

@@

In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly under-appreciated factors that contribute to an open source project's success.

@@

The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient. Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" [@openknowledgefoundation2023]. This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement. However, most open science advocates (including archaeologists, as elicited by @beck2012 and @marwick2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scientific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.

The idea that scientists should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals [@mirowski2018]. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes. However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain [@tennant2020]. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out. Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software.

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scientific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards [cf. @kansa2014; @kintigh2015; @roosevelt2015] distract from fundamental tensions and contradictions regarding the actual value of working in the open. For instance, Faniel et al. [-@faniel2013 299-301], Atici et al. [-@atici2013 676-677], @huggett2018, @sobotkova2018, and @opitz2021 demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate. In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation. Here we aim to identify similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists.

### 2023-09-20 analysis/_03-data_methodology.qmd
Modifications and comments.

<!-- TODO: snappier title? Maybe incoroporating 'open-archaeo'? -->

@@

For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@csardi2023].
This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).

>>>>

<!-- TODO: snappier title? Maybe incorporating 'open-archaeo'? -->

@@

For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@csardi2023].
This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but among those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).

### 2023-09-20 analysis/_04-open_archaeology.qmd
Modifications and comments.

- What kind of projects comprise open archaeology?
  x Breakdown of tags and categories
  x Breakdown of platforms and languages
  x Tags by category
  x Licenses?
- How has this changed over time?
  - Growth of open archaeology by number of commits
  - Growth of open archaeology by number of repos
  - Compare changing proportions of each category on a year by year basis
  - Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology
-->
```
As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists, primarily software but also various forms of open document.
@tbl-categories summarises the categories included.

@@

<!-- TODO: can we say anything more concrete about the tags? -->

>>>>

- What kind of projects comprise open archaeology?
  x Breakdown of tags and categories
  x Breakdown of platforms and languages
  x Tags by category
  x Licenses?
- How has this changed over time?
  - Growth of open archaeology by number of commits
  - Growth of open archaeology by number of repos
  - Compare changing proportions of each category on a year by year basis
  - Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology
-->
```
As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists, primarily software but also various forms of open document.
@tbl-categories summarizes the categories included.

@@

<!-- TODO: can we say anything more concrete about the tags? 
ZB response: We can look at the development of tags over time. For instance, a chart documenting year-over-year growth of each tag based on date of each project's first commit. I imagine this as a stacked bar chart (similar to below) but with segments coded to represent year of first commit. This would fit at the end of this section, under fog-github-cumulative. We can add charts documenting growth of platforms and licenses too. -->

### 2023-09-20 analysis/_05-collaboration.qmd
Modifications and comments.

The span of time between the first and latest commits is visualized more clearly in the following charts. Notably, many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year. Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects. Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects. The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e. the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently. But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates. This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.

>>>>

The span of time between the first and latest commits is visualized more clearly in the following charts. Notably, many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year. Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects. Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects. The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e. the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently. But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterized by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates. This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.

### 2023-09-20 analysis/_06-network.qmd
Modifications and comments.

```{r extract-graph-repo-user, eval=FALSE}
oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
                                   edges_user_to_repo(oarch_graph))
```

```{r fig-graph-repo-user, eval=FALSE}
g_repo_user <- oarch_graph_repo_user |>
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
  geom_node_point(aes(colour = node_type)) +
  scale_edge_alpha(trans = "log")

@@

## repo-repo
We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds. Aside from isolate nodes, which represent repositories with only a single contributor, we detected 21 clusters. While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.

<!-- TODO: Note that we remove disconnected repos? (and how many?) -->

@@

```{r fig-graph-repo-repo, eval=FALSE}
oarch_graph_repo_repo |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  geom_node_point(aes(colour = category)) +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")
```

```{r fig-graph-repo-repo-interactive, eval=FALSE}
iplot <- oarch_graph_repo_repo |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  #geom_node_point(aes(colour = group)) +
  #geom_node_text(aes(label = group)) +
  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")

@@

```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=FALSE}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
                                   edges_common_users(oarch_graph),
                                   directed = TRUE)

@@

ggraph(xx, 'dendrogram', height = height)
```

This graph identifies three core clusters characterized by repositories whose contributors commit to projects other than their own. The clusters that appear along the peiphery largely correspond with the work of single individuals, and sometimes also their colleagues. Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (and perhaps initial) integrations of lone developers into a broader community.

@@

<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph? -->

@@

<!--
 NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include. – JR -->

```{r fig-graph-user-user-centrality, eval=FALSE}
oarch_graph_user_user |> 
  ggraph(layout = "graphopt") +
  # geom_edge_link(aes(alpha = n)) +
  geom_edge_arc(alpha = 0.05) +
  geom_node_point(aes(size = centrality), 
                  shape = 21, colour = "#000000", fill = "#ffffff") +
  # scale_edge_alpha(trans = "log") +
  # labs(edge_alpha = "Common repositories") +
  theme_graph()
```

@@

**NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying theur username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).

<!-- JR: I'm not sure about this. Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study) -->

@@

---

```{r fig-user-betweenness, eval=FALSE}
#| fig-cap: Distribution of betweenness centrality values in the user–user network
oarch_graph_user_user |>
  mutate(
    betweenness = centrality_betweenness(weights = n)
  ) |>
  data.frame() |>
  ggplot(aes(x = betweenness)) +
  geom_density(bw = "nrd") +
  scale_x_log10(labels = scales::number)
```

```{r fig-user-betweenness-activity, eval=FALSE}
#| fig-cap: Correlation between user's centrality (betweenness) in the network and their total number of contributions
oarch_graph_user_user |>
  mutate(
    degree = centrality_degree(), 
    betweenness = centrality_betweenness(weights = n), 
    strength = centrality_degree(weights = n)
  ) |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(betweenness, contribs)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm")
```

>>>>

```{r extract-graph-repo-user, eval=TRUE}
oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
                                   edges_user_to_repo(oarch_graph))
```

```{r fig-graph-repo-user, eval=TRUE}
g_repo_user <- oarch_graph_repo_user |>
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
  geom_node_point(aes(colour = node_type)) +
  scale_edge_alpha(trans = "log")

@@

## repo-repo
We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds. Aside from isolate nodes (n = XXX, not visible in our visualizations), which represent repositories with only a single contributor, we detected 21 clusters. While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.

<!-- TODO: Note that we remove disconnected repos? (and how many?)
ZB response: See change in line 95 above -->

@@

```{r fig-graph-repo-repo, eval=TRUE}
# ZB: I don't think this actually says anything meaningful and could be removed.
oarch_graph_repo_repo |>
  ggraph(layout <- create_layout(oarch_graph_repo_repo, layout = 'igraph', algorithm = 'fr')) +
  geom_edge_link(aes(alpha = 1)) +
  #geom_node_point(aes(colour = category)) +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")
```

```{r fig-graph-repo-repo-interactive, eval=TRUE}
# Remove the legend.
iplot <- oarch_graph_repo_repo |>
  ggraph(layout = "drl") +
  geom_edge_link(aes(alpha = n)) +
  #geom_node_point(aes(colour = group)) +
  #geom_node_text(aes(label = group)) +
  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")

@@

```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=TRUE}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
                                   edges_common_users(oarch_graph),
                                   directed = TRUE)

@@

ggraph(xx, 'dendrogram', height = height)
```
<!--
ZB Note: What we really want out of this is the identification of groups or clusters, as well as the relationships between clusters. In theory, the dendrogram provides the best visualization for this information. I wonder if we can improve it by colour-coding the algorithmically-identified clusters (which should appear side by side), and making the names of the repositories appear vertically aligned perpendicular to the x axis. Additionally, we can colour-code the names of the repositories with their corresponding colours, or indicate that they are part of a cluster using a bracket (like "{" but turned 180 degrees).
-->

This graph identifies three core clusters and several peripheral clusters. The core clusters are characterized by repositories whose contributors commit to projects other than their own. The other peripheral clusters largely correspond with the work of single individuals, and sometimes also their close colleagues. Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.

@@

<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph?
ZB Response: Not much, I think I conceived it more as a means to an end. If I recall correctly, my intention was to ascertain whether different kinds (categories, mainly) of repositories are more central. One observation is the change in magnitude of betweenness centrality between the two "indexical" entries (open-archaeo and ctv-archaeo) and others in this top 10. Perhaps we can detect and visualize additional abrupt "jumps" in magnitude and identify some pattern relating to other variables. -->

@@

<!-- NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include. – JR
ZB Response: I agree, this visual is useless. This, as the other generalized one-mode were meant to simply show that we had indeed split the two-mode into two one-modes. They provide little informational value and should be removed.
-->

```{r fig-graph-user-user-centrality, eval=TRUE}
oarch_graph_user_user |> 
  ggraph(layout = "graphopt") +
  # geom_edge_link(aes(alpha = n)) +
  geom_edge_arc(alpha = 0.05) +
  geom_node_point(aes(size = centrality), 
                  shape = 21, colour = "#000000", fill = "#ffffff") +
  # scale_edge_alpha(trans = "log") +
  # labs(edge_alpha = "Common repositories") +
  theme_graph()
```

@@

**NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying their username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).

<!-- JR: I'm not sure about this. Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study) 
ZB Response: As with the table above, this is a means to an end, a way of working out what we can and want to say about this. However I think this will require more thought, beyond the scope of what we are currently able to achieve, and may be better suited for another paper (connected with the DIIF grant, perhaps). However, see the final note below (line 253) regarding comparison between people who contribute code, things other than code, or both code and non-code, which I think is a viable and valuable addition here.
-->

@@

---
<!--ZB Note: What is this? -->
```{r fig-user-betweenness, eval=TRUE}
#| fig-cap: Distribution of betweenness centrality values in the user–user network
oarch_graph_user_user |>
  mutate(
    betweenness = centrality_betweenness(weights = n)
  ) |>
  data.frame() |>
  ggplot(aes(x = betweenness)) +
  geom_density(bw = "nrd") +
  scale_x_log10(labels = scales::number)
```

```{r fig-user-betweenness-activity, eval=TRUE}
#| fig-cap: Correlation between user's centrality (betweenness) in the network and their total number of contributions
oarch_graph_user_user |>
  mutate(
    degree = centrality_degree(), 
    betweenness = centrality_betweenness(weights = n), 
    strength = centrality_degree(weights = n)
  ) |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(betweenness, contribs)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm")
```

### 2023-09-20 analysis/_07-conclusion.qmd
Modifications and comments.

Archaeology is a "magpie" discipline, using tools and methods that originate from a variety of other disciplines and industries to better understand the past. 
Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-creating with others to get the job done (Batist et al 2021; Morgan and Eve 2012; Caraher 2019). 
It is therefore unsprising that archaeologists are prolific open source software developers.

In this study we have sought to document an emerging community of open source software engineers in archaeology and to assess the extent to which collaboration within this community resembles conventional 'open source' forms. 
We have operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
This excludes a wide range of other potential practices, most obviously the use of other software (though as we demonstrate in section X.X, the git/GitHub combination is overwhelmingly the most popular), but also collaboration through offline or private channels, or forms of collaboration we don't know about. 
We have not directly interviewed those doing the collaborating, though our conclusions do drawn heavily from our experience as members of that community ourselves. 
Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@cite).

>>>>

Archaeology is a "magpie" discipline, using tools and methods that originate from a variety of other disciplines and industries to better understand the past. 
Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-creating with others to get the job done (Batist et al 2021; Morgan and Eve 2012; Caraher 2019). 
It is therefore unsurprising that archaeologists are prolific open source software developers.

In this study, we documented an emerging community of open source software engineers in archaeology and assessed the extent to which collaboration within this community resembles conventional 'open source' forms. 
We operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
This excludes a wide range of other potential practices, most obviously the use of other software (though as we demonstrate in section X.X, the git/GitHub combination is overwhelmingly the most popular), but also collaboration through offline or private channels, or forms of collaboration we don't know about. 
We have not directly interviewed those doing the collaborating, though our conclusions do draw heavily from our experience as members of that community ourselves. 
Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@ducke2013 10.4324/9781315431932-12; @whallon1972 10.1007/BF02403759).
